---
title: "DM-Q5-tp2"
author: "Mohamed Moukhtar / KHALIL"
date: "2023-04-08"
output:
  html_document: default
  word_document: default
lang: fr
---

## Data Importation:

```{r}
Demo <- read.csv("C:/Users/USER/Downloads/Demo_csv (1).csv",sep=';')
head(Demo)
```

Produce a numerical summary of the variable "carcat":

```{r}
table(Demo$carcat)
```

Produce a numerical summary of the variable "carcat":

```{r}
library(ggplot2)
ggplot(Demo, aes(x = carcat, fill = carcat)) +
  geom_bar(aes(y = ..count..), width = 0.5) +
  geom_text(aes(label = ..count..), stat = "count", 
            position = position_stack(vjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
ggplot(Demo, aes(x = "", fill = carcat)) +
  geom_bar(width = 1) +
  stat_count(aes(label = paste0(scales::percent(..count../sum(..count..)), " (", ..count.., ")")),
             geom = "text", position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(fill = "La categorie des voitures")

```

Create the variable "carcatNA" with 30% missing values:

```{r}
set.seed(123)
sample_indices <- sample(1:nrow(Demo), round(0.3*nrow(Demo)), replace=FALSE)
carcatNA <- Demo$carcat
carcatNA[sample_indices]=NA
Demo$carcatNA = carcatNA
```

Cross the variable "carcat" with "carcatNA":

```{r}

addmargins(table(Demo$carcat,Demo$carcatNA,useNA = 'ifany'))
addmargins(table(Demo$carcat[sample_indices],Demo$carcatNA[sample_indices],useNA = 'ifany'))
```

## Imputation: Simple Hot-deck
```{r}
library(VIM)
Demo_hd=hotdeck(Demo)
Demo$carcat_hds=Demo_hd$carcatNA
addmargins(table(Demo$carcat[sample_indices],Demo$carcat_hds[sample_indices],useNA = "ifany"))


```
Let's look at the accuracy score:
```{r}

accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcat_hds[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
print(accuracy)
```
## Imputation: Conditional Hot-deck

```{r}
df_imp <- hotdeck(Demo, domain_var = 'income')
Demo$carcat_hdc=df_imp$carcatNA
addmargins(table(Demo$carcat[sample_indices],Demo$carcat_hdc[sample_indices]))


```
Let's look at the accuracy score for conditional hot-deck:
```{r}

accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcat_hdc[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
print(accuracy)
```

## Imputation: Simple Mode


```{r}
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
}
mode = getmode(Demo$carcat)

carcat_ms = Demo$carcatNA
Demo$carcat_ms <- ifelse(is.na(carcat_ms),mode,carcat_ms)
```

Output the confusion matrix and accuracy:

```{r}

addmargins(table(Demo$carcat[sample_indices],Demo$carcat_ms[sample_indices]))


accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcat_ms[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
print(accuracy)

```
## Imputation: Conditional Mode

Create a list that contains the mode of the car category variable given the income category:
```{r}
Demo$carcat_mc = Demo$carcatNA
# Get unique modalities of inccat
inc_levels <- unique(Demo$inccat)
  
  # Initialize the dictionary for conditional modes
mode_cond <- list()
  
  # Iterate over inccat modalities and calculate the mode of carcat
for (inc_level in inc_levels) {
    # Filter data by inccat modality
    Demo_subset <- Demo[Demo$inccat == inc_level,]
    
    # Calculate the mode of carcat
    car_mode <- names(sort(-table(Demo_subset$carcat)))[1]
    
    # Add the mode to the dictionary
    mode_cond[[inc_level]] <- car_mode
}
print(mode_cond)



```

Perform imputation on the variable "carcat_mc":
```{r}

  # Filtrer les données avec des valeurs manquantes pour carcat
missing_Demo <- Demo[is.na(Demo$carcat_mc),]
  
  # Imputer les valeurs manquantes avec les modes conditionnels
for (i in 1:nrow(missing_Demo)) {
    inc_level <- missing_Demo$inccat[i]
    Demo$carcat_mc[is.na(Demo$carcatNA) & Demo$inccat == inc_level] <- mode_cond[[inc_level]]
}

```
Evaluate this imputation using the confusion matrix and accuracy score:

```{r}
addmargins(table(Demo$carcat[sample_indices],Demo$carcat_mc[sample_indices]))

accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcat_mc[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
print(accuracy)

```


## Imputation: FDA

```{r}

nums <- unlist(lapply(Demo, is.numeric), use.names = FALSE)  
data=Demo[ , nums]
data$groupe=Demo$carcatNA
base=na.omit(data)

library(MASS)
res <- lda(groupe ~.,base)
res

coef=as.matrix(res$scaling,nrow=1,ncol=6)

for (i in 1:dim(base)[1]) {
  if((base$groupe[i]) == "Economique"){
    score1=sum(t(coef)*as.numeric(as.matrix(base[i,1:5])))}
  if((base$groupe[i]) == "Luxe"){
    score2=sum(t(coef)*as.numeric(as.matrix(base[i,1:5])))}
  if((base$groupe[i]) == "Standard"){
    score3=sum(t(coef)*as.numeric(as.matrix(base[i,1:5])))}
  
}


gain <- rep(NA, dim(data)[1])
Demo$carcatAFD <- Demo$carcatNA

for (i in 1:dim(data)[1]) {
  if (is.na(Demo$carcatNA[i]) == TRUE) {
    gain[i] <- sum(t(coef) * as.numeric(as.matrix(data[i, 1:5])))
    s1 <- abs(gain[i] - score1)
    s2 <- abs(gain[i] - score2)
    s3 <- abs(gain[i] - score3)
    
    if (s1 < s2 & s1 < s3 ) {
      Demo$carcatAFD[i] <- "Economique"
    } else if (s2 < s1 & s2 < s3 ) {
      Demo$carcatAFD[i] <- "Luxe"
    } else if (s3 < s1 & s3 < s2 ) {
      Demo$carcatAFD[i] <- "Standard"
    } 
  }
}

```

Evaluate this imputation:
```{r}

addmargins(table(Demo$carcat[sample_indices],Demo$carcatAFD[sample_indices]))


accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcatAFD[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
print(accuracy)


```

## Imputation :logit-probit

```{r}

library(mice)

# Subset the relevant columns
df_logit <- Demo[c("age", "inccat", "carcatNA")]

# Recode carcatNA to 0/1
df_logit$carcatNA <- ifelse(df_logit$carcatNA == "Economique",0,ifelse(df_logit$carcatNA == "Luxe", 0.5, 1))

# Create a formula for logistic regression
formula_logit <- as.formula("carcatNA ~ age + inccat")

# Create a method object for logistic regression
meth_logit <- "logreg"

# Set the number of imputations
n_imp <- 5

# Impute missing values using mice
imputed_logit <- mice(df_logit, method = meth_logit, formulas = list(formula_logit), m = n_imp)

# Complete the imputation
df_logit_imputed <- complete(imputed_logit)

df_logit_imputed$carcatNA <- ifelse(df_logit_imputed$carcatNA == 0, 'Economique', ifelse(df_logit_imputed$carcatNA == 0.5,'Luxe','Standard'))
Demo$carcat_logit = df_logit_imputed$carcatNA



```

Evaluate this imputation method:
```{r}
addmargins(table(Demo$carcat[sample_indices],Demo$carcat_logit[sample_indices]))

accuracy <- sum(Demo$carcat[sample_indices] == Demo$carcat_logit[sample_indices], na.rm = TRUE) / sum(!is.na(Demo$carcat[sample_indices]))
accuracy
```



```